---
title: "LGCM_MEPRS"
author: "P. Hoover"
date: "10/26/2021"
output: 
  html_document:
    keep_md: TRUE
    includes:
      in_header: test.html
    theme: flatly
    highlight: tango
---

Below outlines the expansion of the LGCM with the use of MEPRS3. We will attempt to identify and evaluate latent groups among varying healthcare utilization trajectories. 

# Methodology

## Data Sources:

* Direct Care - Inpatient/Outpatient
* Purchase Care - Inpatient/Outpatient

## Sample Population:

1. Identify those with an initial mTBI within DoD (DC/PC/IN/OUT)
2. Verify AD status on index date
3. Verify no more severe dx of TBI within year following index date
4. Verify HC utilization < -365 & > 365 days since index date (actively using system)

## Analysis:

1. Extract all DC Outpatient encounters 1 year following index mTBI
2. Map MEPRS 3 to clinic categories (e.g., MH, Neurology/Pain, Physical/Rehab, etc.)

  * Removed any A-Meprs (Inpatient)
  
3. For each month following index mTBI, sum total number of encounters specific to each clinic
4. EDA on current clinic and sample population
5. Conduct Latent Growth Curve Modeling
6. Conduct Post-hoc analysis - differences among latent groups 


```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
#Libraries
library(tidyverse)
library(flexmix)
library(ggplot2)
library(kableExtra)

#Functions
'%ni%'<-Negate('%in%')

grp_prp<-function(df,x){
  x <- enquo(x)
  df %>%
    group_by(!!x)%>%
    summarise(n=n())%>%
    mutate(Prop=round(n/sum(n)*100,1))%>%
    knitr::kable()%>%
    kableExtra::kable_styling(full_width = F)
}

grp_prp_2<-function(df,x,y){
  xx <- enquo(x)
  yy <- enquo(y)
  test<-df %>%
    group_by(!!xx,!!yy)%>%
    summarise(n=n())%>%
    mutate(Prop=round(n/sum(n)*100,1))%>%
    mutate(final=paste(n," (",Prop,")",sep=""))%>%
    select(-n,-Prop)

  test%>%
   pivot_wider(id_cols=!!xx,names_from=!!yy,values_from=final)%>%
    knitr::kable()%>%
    kableExtra::kable_styling(full_width = F)
}

#################
### Load Data ###
#################

#Pt List 
pts<-read.csv("/Users/peterhoover/Documents/Analysis_Projects/R_Scripts/LGCV_Clinics/Data/Clean/pts_filtered.csv")
pts<-pts%>%
  mutate(age = case_when(pseudo_personid==1262865906 ~ 44.13,
                         TRUE ~ age))%>%
  mutate(age_grp = cut(floor(age),breaks=c(0,24,34,44,54,100),
                       labels=c("17to24","25to34","35to44","45to54","55+"),include.lowest = TRUE))%>%
  mutate(sponservice_updated = case_when(sponservice %in% c("Air Force","Army","Marines","Navy")~paste(sponservice),
                                 TRUE ~ paste("Other_Unknown")))%>%
  mutate(rank_class_updated = case_when(rank_class=="Unknown"|is.na(rank_class)~"Other_Unknown",
                                TRUE ~ paste(rank_class)))

#pts%>%group_by(age_grp)%>%summarise(min=min(age),max=max(age))

#Clinic Cnts
clin_long<-read.csv("/Users/peterhoover/Documents/Analysis_Projects/R_Scripts/LGCV_Clinics/Data/Clean/clinic_long.csv")

#Encs
encs<-read.csv("/Users/peterhoover/Documents/Analysis_Projects/R_Scripts/LGCV_Clinics/Data/Clean/enc_clean.csv")
```

# Demographics

Sample Population: `r length(unique(pts$pseudo_personid))`

Gender: `r grp_prp(pts,gender)`

Age: `r grp_prp(pts,age_grp)`

Ethnicity: `r grp_prp(pts,ethnicity)`

Service: `r grp_prp(pts,sponservice_updated)`

Rank: `r grp_prp(pts,rank_class_updated)`

# Healthcare Utilization

```{r}
cnts<-encs%>%
  group_by(pseudo_personid)%>%
  summarise(n=n())%>%
  right_join(pts)%>%
  mutate_at(vars(n),~replace(.,is.na(.),0))

clin_sum<-clin_long%>%
  group_by(pseudo_personid,Clinic)%>%
  summarise(n=sum(n))%>%
  ungroup()%>%
  group_by(Clinic)%>%
  summarise(mean=round(mean(n),2),med=median(n))%>%
    knitr::kable()%>%
    kableExtra::kable_styling(full_width = F)

clin_map<-encs%>%group_by(meprs3,Description,Clinic)%>%summarise(n=n())%>%
  DT::datatable()

```

Total Encounters: `r nrow(encs)`

Average (Med) Encounter: `r paste(round(mean(cnts$n),2)," (",median(cnts$n),")")`

Clinic Summary: `r clin_sum`

Clinic Mappings: `r clin_map`







