---
title: "Latent Growth Curve Models"
author: "P. Hoover"
date: "Apri 1, 2021"
output: 
  html_document:
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


```{r}
#Libraries
library(tidyverse)
library(flexmix)
library(ggplot2)
library(kableExtra)

#Previous script
source("PreProcess.R")

#Data
#Patient with tbi 2016-2017, initial mtbi, no susbsequent tbi, no person hx tbi, and HC utilization > and < 365 from index date 
pts<-data.table::fread("/Users/peterhoover/Documents/Analysis_Projects/Global_data/cost/clean/perPatientTotals_proj.csv")

#Load previous model
load("nb_model1.rda")

#Old class assignments with sample patients 
class_ass<-read.csv(file="class_assignments.csv")
```

# Sample Dataset

1. First diagnosis of mTBI

2. Verify no more severe dx within year following

3. Verify health care utilization 

4. Obtain all DC encounters within 1 year either side of mTBI 

## Objective

* Evaluate changes in health care utilization as a result of the mTBI 

  + Overall changes in total encounters

  + Changes in type of encounters

  + Identify high utilizers 


# Examine Change in Encounters {.tabset}

## Frequency 

* Increase directly following mTBI and a gradual reduction

```{r}
ggplot(encs,aes(x=daydiff))+geom_histogram(bins=50)+ggtitle("Histogram of Encounters wrt Index Date")
```

## Overall

* 86.54% Increase in overall utilization

* Increase in 7.7 average encounters, median increase of 4

```{r}
encs<-encs%>%
  mutate(enc_flag=factor(enc_flag,ordered=T,levels=c("Pre","Post")))

enc_cnts<-encs%>%
  group_by(enc_flag)%>%
  summarise(n=n())

enc_sum<-pts%>%
  select(pseudo_personid,n_Post,n_Pre)%>%
  reshape2::melt(id.var="pseudo_personid")%>%
  group_by(variable)%>%
  summarise(mean=mean(value),med=median(value))%>%
  mutate(variable=factor(variable,levels=c("n_Pre","n_Post"),ordered=T))

gridExtra::grid.arrange(ggplot(enc_cnts,aes(x=enc_flag,y=n))+geom_bar(stat='identity')+ggtitle("Total Number of Encounters"),
ggplot(enc_sum,aes(x=variable,y=mean))+geom_bar(stat='identity')+ggtitle("Mean (Med) Number of Encounters")+geom_point(aes(y=med)),ncol=2)
```

# Following Injury Summarization {.tabset}

## Encounter Overall

```{r}
mon_sum<-encs%>%
  filter(enc_flag=="Post")%>%
  mutate(Mon_Flag_raw = daydiff/30)%>%
  mutate(Mon_Flag = floor(Mon_Flag_raw))%>%
  #mutate(Mon_Flag = ceiling(Mon_Flag_raw))%>%
  filter(Mon_Flag<12)%>%
  group_by(pseudo_personid,Mon_Flag)%>%
  summarise(n=n())

mon_sum_cnts<-expand.grid(pseudo_personid=unique(pts$pseudo_personid),Mon_Flag=unique(mon_sum$Mon_Flag))

mon_sum_cnts<-left_join(mon_sum_cnts,mon_sum,by=c("pseudo_personid","Mon_Flag"))

mon_sum_cnts$n[is.na(mon_sum_cnts$n)]<-0

mon_sum_sums<-mon_sum_cnts%>%
  group_by(Mon_Flag)%>%
  summarise(mean=mean(n),med=median(n))

ggplot(mon_sum_sums,aes(x=Mon_Flag,y=mean))+geom_bar(stat='identity')+ggtitle("Mean (Med) Number of Encounters")+geom_point(aes(y=med))
```

## Encounters per Clinic

```{r}
clinic_sums<-clinic_cnts%>%
  group_by(Mon_Flag,prodline)%>%
  summarise(mean=mean(n),med=median(n))

ggplot(clinic_sums,aes(x=Mon_Flag,y=mean))+geom_bar(stat='identity')+ggtitle("Mean (Med) Number of Encounters")+geom_point(aes(y=med))+facet_wrap(~prodline)
```

# Summary

* Population level, provides nice overview of how encounters change 

  + What/Where are the clinic demands?

  + What can we expect patients to utilize?
    
   + e.g., Expect the average patient to have 20.4 encounters following injury

* This also assumes that all patients are generally equal

  + Subset of patients that are high utilizers of MH services


***

# Latent Growth Curve Models

* Longitudinal analysis technique to estimate growth over a period of time

* Frequently used in social sciences, examining changes in traits

* Considers change time in terms of an underlying, latent process

* More flexible than multilevel modeling or repeated ANOVA

  + Inclusion of additional predictors (e.g., demographics, types of outcomes)


## Objective

+ Identify latent groups of HC utilization within the year following mTBI

+ For each patient, calculated total number of encounters per ProdLine (clinic) each month following injury


## LGCM - Procedures

1. Evaluate distribution fit
  + Gaussian, Poisson, NB, Zero-Inflated, etc.

2. Determine growth parameters
  + Linear, polynomial

3. Estimate latent groups
  + k latent groups (2-7)

  + E-M: For each k, iteratively obtain parameter estimates and posterior estimates of patient class membership in each k-group to maximize likelihood of the data give the model parameters 
  
  + Obtain solution for best fitting parameters to meet the number of estimated latent groups 

4. Further include other predictors (demographics, etc.)


## Model Summaries

```{r}
nb_model1
```
# Model Estimates

```{r}
getModel(nb_model1)@components

```

# Evaluate Number of Patients per Group

```{r}
set.seed(111)
sam<-sample(unique(clinic_cnts$pseudo_personid),1000) #Sample population for testing 
sample_df<-clinic_cnts%>%
  filter(pseudo_personid %in% sam)

#sample_df<-clinic_cnts #whole population

sample_df$Mon_Flag<-as.numeric(sample_df$Mon_Flag)

class_ass<-cbind(sample_df,getModel(nb_model1)@cluster)

class_ass%>%
  distinct(pseudo_personid,`getModel(nb_model1)@cluster`)%>%
  group_by(`getModel(nb_model1)@cluster`)%>%
  summarise(n=n())%>%
  mutate(prop=n/sum(n)*100)%>%
  kable()%>%kable_styling(full_width=F)
```

# Explore How Utilization Changes 

```{r,fig.width = 10, fig.height = 10}
names(class_ass)[5]<-"cluster"
class_sum<-class_ass%>%
  ungroup()%>%
  group_by(prodline,Mon_Flag,cluster)%>%

  summarise(min=min(n),max=max(n),mean=mean(n),med=median(n))

ggplot(class_sum%>%filter(cluster==1),aes(x=Mon_Flag,y=mean))+
  geom_point(aes(color=as.factor(cluster)))+facet_wrap(~prodline)+
  geom_line(aes(color=as.factor(cluster)))

# ggplot(class_sum,aes(x=Mon_Flag,y=med))+
#   geom_point(aes(color=as.factor(cluster)))+facet_wrap(~prodline)+
#   geom_line(aes(color=as.factor(cluster)))

```

# Compare Demographics {.tabset}

## Gender

```{r}
class_distinct<-class_ass%>%
  distinct(pseudo_personid,cluster)

pts_final<-pts%>%
  select(pseudo_personid,gender,ethnicity,age,sponservice)%>%
  inner_join(class_distinct,by="pseudo_personid")%>%
  mutate(sponservice=case_when(sponservice==""~"X",
                               TRUE ~ paste(sponservice)))

#Gender
pts_final%>%
  group_by(cluster,gender)%>%
  summarise(n=n())%>%
  mutate(Prop=round(n/sum(n)*100,2))%>%
  reshape2::dcast(gender~cluster,value.var="Prop")%>%View()
  kable()%>%kable_styling(full_width=F)

#prop.test(table(pts_final$cluster,pts_final$gender))
```

## Age

```{r}
#Age

pts_final%>%
  group_by(cluster)%>%
  summarise(min=round(min(age),2),mean=round(mean(age),2),max=round(max(age),1))%>%View()
  kable()%>%kable_styling(full_width=F)

#aov_1<-aov(age~as.factor(cluster),data=pts_final)
#summary(aov_1)
#TukeyHSD(aov_1)
```

## Race

```{r}
#Ethnicity

pts_final%>%
  group_by(cluster,ethnicity)%>%
  summarise(n=n())%>%
  ungroup()%>%
  group_by(cluster)%>%
  #group_by(ethnicity)%>%
  mutate(Prop=round(n/sum(n)*100,2))%>%
  reshape2::dcast(ethnicity~cluster,value.var="Prop")%>%
  kable()%>%kable_styling(full_width=F)

#prop.test(table(pts_final$cluster,pts_final$ethnicity))

```

## BoS

```{r}
#BoS

pts_final%>%
  group_by(cluster,sponservice)%>%
  summarise(n=n())%>%
  ungroup()%>%
  group_by(cluster)%>%
  #group_by(sponservice)%>%
  mutate(Prop=round(n/sum(n)*100,2))%>%
  reshape2::dcast(sponservice~cluster,value.var="Prop")%>%
  kable()%>%kable_styling(full_width=F)

```

